Inferring the effective start dates of non-pharmaceutical interventions during \covid outbreaks

### 2020-08-05 ms/ms.tex
sig statement; orcid

\affil[1]{School of Computer Science, Interdisciplinary Center Herzliya, Herzliya 4610101, Israel}
\affil[2]{School of Public Health, Tel Aviv University, Tel Aviv 6997801, Israel}{
\affil[3]{Porter School of the Environment and Earth Sciences, Tel Aviv University, Tel Aviv 6997801, Israel}
\affil[a]{Corresponding author: yoav@yoavram.com}

@@

@@

>>>>

\affil[1]{School of Computer Science, Interdisciplinary Center Herzliya, Herzliya 4610101, Israel}
\affil[2]{School of Public Health, Tel Aviv University, Tel Aviv 6997801, Israel}{
\affil[3]{Porter School of the Environment and Earth Sciences, Tel Aviv University, Tel Aviv 6997801, Israel}
\affil[a]{Corresponding author: yoav@yoavram.com, ORCID 0000-0002-9653-4458}

@@

Keywords: SEIR, COVID-19, public health, pandemic, infectious disease

Classification: Biological Sciences / Population Biology; Social Sciences / Environmental Sciences

%\pagebreak

@@

\renewcommand{\abstractname}{Significance statement}
\begin{abstract}
We have inferred the effective start date of non-pharmaceutical interventions, like lockdowns, during the initial \covid outbreak by fitting an epidemiological model to daily case data in 12 countries. Our results show that the effective start of interventions in many cases occurs 3-10 days later then the official day. Moreover, we demonstrate that such delays can lead to under-estimation of the intervention impact, and thus cause decision makers to implement stricter interventions. We suggest that a late effective start of interventions may be the result of low compliance to authority or conformity to social norms. Our study therefore highlights the importance of further research into the role of human behavior and social organization on the current and future epidemics.
\end{abstract}
